---
title: "Associative Analysis-"
author: "Karen Mwaura"
date: "2/4/2022"
output: html_document
---


```{r}
# Loading the arules library
#
library(arules)

```


```{r}
library(readr)
df<- read_csv("C:/Users/user/Downloads/Supermarket_Sales_Dataset II.csv")
head(df, 5)

```

```{r}
# Loading our transactions dataset from our csv file
path <- "C:/Users/user/Downloads/Supermarket_Sales_Dataset II.csv"

Transactions<-read.transactions(path, sep = ",")
Transactions


```


```{r}
# Verifying the object's class

class(Transactions)

```

```{r}
# Previewing our first 5 transactions

inspect(Transactions[1:5])

```

```{r}
# If we wanted to preview the items that make up our dataset,
# alternatively we can do the following..

items<-as.data.frame(itemLabels(Transactions))
colnames(items) <- "Item"
head(items, 10)    

```


```{r}
#Get summary of the transaction dataset
# This would give us some information such as the most purchased items

summary(Transactions)

```

```{r}
# Exploring the frequency of some articles 

itemFrequency(Transactions[, 8:10],type = "absolute")
round(itemFrequency(Transactions[, 8:10],type = "relative")*100,2)

```

```{r}
# Displaying top 10 most common items in the transactions dataset 
# and the items whose relative importance is at least 10%

par(mfrow = c(1, 2))

```

```{r}

# plot the frequency of items
itemFrequencyPlot(Transactions, topN = 10,col="purple")
itemFrequencyPlot(Transactions, support = 0.1,col="sky blue")
```

```{r}
# Building a model based on association rules using the apriori function

rules <- apriori (Transactions, parameter = list(supp = 0.001, conf = 0.8))
rules

```

```{r}
# Building a apriori model with Min Support as 0.002 and confidence as 0.8.

rules2 <- apriori (Transactions,parameter = list(supp = 0.002, conf = 0.8)) 

```


```{r}
# Building apriori model with Min Support as 0.002 and confidence as 0.6.

rules3 <- apriori (Transactions, parameter = list(supp = 0.002, conf = 0.6)) 

```

There's an increase in the minimum support of 0.001 to 0.002 and model rules went from 74 to 2. This means that, a high level of support can make teh model lose interesting rules. The other part there was a decrease in the minimum confidence to 0.6 and the model rules went from 2 to 43. This means, loew confidence level will increase teh number of rules of which many may not be useful.

```{r}
# perform an exploration of our model 

summary(rules)

```

```{r}
# Observing rules built in our model 

inspect(rules[1:5])

```

If one buys frozen smoothie with spinach, they are 89% likely to buy mineral water.
If a person buys bacon and pancakes, they're 81% likely to buy spaghetti.
If one buys turkey and nonfat milk, they're 81% likely to get mineral water.
If one gets groundbeef and nonfat milk, they have 85% chance of buying mineral water.
If one buys mushroom cream sauce and pasta, they are 95% likely to buy escalope.

```{r}
# Ordering these rules by a criteria such as the level of confidence then looking at the first five rules.

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])

```

The first four rules have a confidence of 100 and the last oen has 95.

```{r}
# If we wanted to make a promotion relating to sale of milk,then create a subset of rules in regards to these products 
# This would tell us the items that the customers bought before purchasing milk.

milk <- subset(rules, subset = rhs %pin% "milk")

# Then order by confidence
milk<-sort(milk, by="confidence", decreasing=TRUE)
milk

inspect(milk[1:5])

```

```{r}
# We want to determine items that customers might buy who have previously bought milk

# Subset the rules
milk <- subset(rules, subset = lhs %pin% "milk")

# Order by confidence
milk<-sort(milk, by="confidence", decreasing=TRUE)

# inspect top 5
inspect(milk[5:15])

```





